Critical Temperature of the Ising Ferromagnet on Proximity Graphs derived from Square Lattices by Site Displacement

### 2013-12-02 Tex/paper/model.tex
Proximity graph abschnitt

% TODO: alles hierdrunter
\subsection{Proximity Graphs}
\label{ssec:graphtypes}
    A graph \(G(V,E)\) is a set of nodes \(V\) and edges \(E \subset V^{(2)}\). In the
    scope of this model, the nodes get identified with the sites of the
    lattice and the edges signal the neighbor status of two sites. If the
    corresponding nodes are connected, the sites are neighbors.\\
    All here mentioned graph types are \emph{proximity graphs}.
    The edges of these graphs connect nodes which are by some metric near
    to each other.
    Hence they are suited to generalize problems defined on regular
    lattices with nearest neighbor relationships.
    In this thesis the distance is always determined by the Euclidean
    metric in two dimensions, though in principle every metric in any
    dimension can be used.\\

    \subsubsection{Delaunay Triangulation}
        The Delaunay triangulation (DT) is an undirected graph. An edge
        between two nodes \(i\) and \(j\) will be drawn, if there exists
        a circle passing through \(i\) and \(j\), which does not contain
        any other node in its interior, see Ref.\ \cite{Katajainen}.

    \subsubsection{Gabriel Graph}
        The Gabriel graph (GG) \cite{Gabriel1969}
        is a subgraph of the DT, i.e.\ for the same set of nodes
        \(V\) the edge set of the DT is a superset of the edge set of the
        GG \(E_{DT} \supseteq E_{GG}\). Two nodes \(i\) and \(j\) with distance
        \(d_{ij}\) will be connected with an edge, if a circle with its
        center on half way between \(i\) and \(j\) and radius
        \(r = \frac d 2\) contains no other nodes. This area will be
        called \emph{lune} in the following. See also the cross hatched region
        from Fig.\ \ref{fig:lunes}.

    \subsubsection{Relative Neighborhood Graph}
        The Relative Neighborhood
        graph (RNG) \cite{Toussaint1980} is a subgraph of the GG. Two nodes \(i\) and \(j\) with
        distance \(d_{ij}\) will be connected again, if no other node is in the
        \emph{lune}, which is different from the lune of the GG. The lune
        is defined as the intersection of two
        circles with radius \(r = d\) and centers on \(i\) and \(j\).
        See also the hatched region in Fig.\ \ref{fig:lunes}.

    \begin{figure}[htbp]
        \centering
        \includegraphics[width=0.15\textwidth]{images/GG_RNG_def}
        \includegraphics[width=0.15\textwidth]{images/RNG/L12S03.pdf}
        \includegraphics[width=0.15\textwidth]{images/GG/L12S03.pdf}
        \caption[Gabriel- and Relative Neighborhood Graph]
        {
            Left: Lunes, which define whether an edge
            exist, of RNG (hatched region) and GG (cross hatched region).
            It is evident from this sketch that the GG is a supergraph
            of the RNG. If there is an edge in the RNG, the hatched
            region contains no node, then of course also the double
            hatched region contains no node and thus this edge appears
            also in the GG. So every egde of the RNG is also present
            in a GG on the same set of nodes.
            Example of a RNG (center) and GG (right) on periodic boundary conditions.
            Periodic nodes are dashed.
        }
        \label{fig:lunes}
    \end{figure}

    \subsubsection{Construction}
        In this thesis only GG and RNG are examined.
        The simplest way to construct these graphs is to test for each
        pair of nodes if any other node lies in
        the lune of the pair. That is of complexity \(O (N^3)\), because
        there are \(N(N-1)\) pairs and for each \((N-2)\) nodes to test. So
        the running time of a straight forward implementation would be of
        order \(O(N^3)\).\\
        To reduce the complexity one can first create a DT in complexity \(O (N \log N)\)
        \cite{RNGCell} and test the connectedness for each edge contained
        therein in \(O (N)\), because the DT is a supergraph of both.
        This results in a complexity of \(O (N^2 \log N)\)
        For the construction of the DT for a given set of points one might
        use existing software libraries, e.g.\ the \texttt{qHull}\cite{qhull} library.
        However, generation of the graphs is not time critical in the scope
        of this bachelor thesis.\\
        So a trade off is to use basically the simple method but only test
        the connectedness for nodes which are near to the lune and abort if
        one node inside the lune is found. To determine which nodes are
        near the lune one can subdivide the area in \(M\) \emph{cells} of size
        \(L_c \times L_c\) and save for each cell a list with nodes lying
        inside like presented in Ref.\ \cite{RNGCell} and sketched in Fig.\ \ref{fig:cell}.
        The implementation of this thesis uses \(M = N\) (\(L_c = 1\)).
        \begin{figure}[htbp]
            \centering
            \input{images/CellBig}
            \input{images/CellSmall}
            \caption[Sketch how the Cell Method works]
            {
                Sketch how the cell algorithm for the construction of the
                RNG and GG works. Here with the lune definition of the GG.
                The bounding box of the lune, which determines which cells have
                to be tested, is marked with thick colored lines.
                Left: shows that it is sufficient to find a
                single node in an inner cell of the bounding box to discard
                the edge between two distant nodes. The nodes inside the
                other 8 marked cells do not have to be tested anymore.
                Right: shows that only the nodes inside the
                marked cell have to be tested, because there are no other nodes,
                the edge can be drawn.
            }
            \label{fig:cell}
        \end{figure}\\
        Given the discretized cell structure it is just necessary to test the nodes in the cells which
        resemble a rectangular bounding box of the lune. Most pairs will be
        far away from each other and there will be one or more cells in the
        middle of the bounding box, which are completely inside the lune.
        Then only one node has to be tested from such a cell to discard an
        edge between the nodes. On the other hand, nodes that will be connected
        are near to each other so that only very few cells intersect the bounding
        box of their lune and consequently only very few nodes have to be tested.\\
        Here the number of cells equals the number of nodes.
        Indeed this method reduced the time needed to construct a RNG with
        \(N=32^2\) and \(N=64^2\) by a factor of
        over \(15\) respectively \(40\). Though the complexity is still of
        order \(O(N^2)\) in the best case, because for every pair at least
        one check has to be performed.

>>>>

\subsection{Proximity Graphs}
\label{ssec:graphtypes}
    A graph \(G(V,E)\) is a set of nodes \(V\) and edges \(E \subset V^{2}\).
    In the scope of this model, the nodes get identified with the sites of the
    lattice. If two nodes are connected by an edge, the sites are neighbors.\\
    All here mentioned graph types are \emph{proximity graphs}.
    The edges of these graphs connect nodes which are by some metric near
    to each other.
    Hence they are suited to generalize problems defined on regular
    lattices with nearest neighbor relationships.
    To determine the distance we use the Euclidean
    metric on two-dimensional graphs, though in principle every metric in any
    dimension can be used.\\

    \subsubsection{Gabriel Graph}
        The Gabriel graph (GG) \cite{Gabriel1969}
        is a subgraph of the Delaunay triangulation (DT) (for a definition
        see e.g.\ Ref.\ \cite{Katajainen}), i.e.\ for the same set of nodes
        \(V\) the edge set of the DT is a superset of the edge set of the
        GG \(E_{DT} \supseteq E_{GG}\). Two nodes \(i\) and \(j\) with distance
        \(d_{ij}\) will be connected with an edge, if a circle with its
        center on half way between \(i\) and \(j\) and radius
        \(r = \frac d 2\) contains no other nodes. This area will be
        called \emph{lune} in the following. See also the cross hatched region
        from Fig.\ \ref{fig:lunes}\subref{sfig:lunes:def}. An Example is
        given in Fig.\ ref{fig:lunes}\subref{sfig:lunes:gg}.

    \subsubsection{Relative Neighborhood Graph}
        The Relative Neighborhood
        graph (RNG) \cite{Toussaint1980} is a subgraph of the GG. Two nodes \(i\) and \(j\) with
        distance \(d_{ij}\) will be connected again, if no other node is in the
        \emph{lune}, which is different from the lune of the GG. The lune
        is defined as the intersection of two
        circles with radius \(r = d\) and centers on \(i\) and \(j\).
        See also the hatched region in Fig.\ \ref{fig:lunes}. An Example is
        given in Fig.\ ref{fig:lunes}\subref{sfig:lunes:rng}.

    \begin{figure}[htbp]
        \centering
        \subfigure[Definition of the Lunes][]{
            \label{sfig:lunes:def}
            \includegraphics[width=0.14\textwidth]{images/GG_RNG_def}
        }
        \subfigure[RNG example][]{
            \label{sfig:lunes:rng}
            \includegraphics[width=0.14\textwidth]{images/RNG/L12S03.pdf}
        }
        \subfigure[GG example][]{
            \label{sfig:lunes:gg}
            \includegraphics[width=0.14\textwidth]{images/GG/L12S03.pdf}
        }
        \caption[Gabriel - and Relative Neighborhood Graph]
        {
            \subref{sfig:lunes:def} Lunes, which define whether an edge
                exist, of RNG (hatched region) and GG (cross hatched region).
                It is evident from this sketch that the GG is a supergraph
                of the RNG. If there is an edge in the RNG, the hatched
                region contains no node, then of course also the double
                hatched region contains no node and thus this edge appears
                also in the GG. So every egde of the RNG is also present
                in a GG on the same set of nodes.
            \subref{sfig:lunes:rng} Example of a RNG on periodic
                boundary conditions.
            \subref{sfig:lunes:gg} Example of a GG on
                periodic boundary conditions. Periodic nodes are dashed.
        }
        \label{fig:lunes}
    \end{figure}

    % TODO: alles hierdrunter
    \subsubsection{Construction}
        The simplest way to construct these graphs is to test for each
        pair of nodes if any other node lies in
        the lune of the pair. That is of complexity \(O (N^3)\), because
        there are \(N(N-1)\) pairs and for each \((N-2)\) nodes to test. So
        the running time of a straight forward implementation would be of
        order \(O(N^3)\).\\
        To reduce the complexity one can first create a DT in complexity \(O (N \log N)\)
        \cite{RNGCell} and test the connectedness for each edge contained
        therein in \(O (N)\), because the DT is a supergraph of both.
        This results in a complexity of \(O (N^2 \log N)\)
        For the construction of the DT for a given set of points one might
        use existing software libraries, e.g.\ the \texttt{qHull}\cite{qhull} library.\\
        So a trade off is to use basically the simple method but only test
        the connectedness for nodes which are near to the lune and abort if
        one node inside the lune is found. To determine which nodes are
        near the lune one can subdivide the area in \(M\) \emph{cells} of size
        \(L_c \times L_c\) and save for each cell a list with nodes lying
        inside like presented in Ref.\ \cite{RNGCell} and sketched in Fig.\ \ref{fig:cell}.
        The implementation of this thesis uses \(M = N\) (\(L_c = 1\)).
        \begin{figure}[htbp]
            \centering
            %~ \input{images/CellBig}
            %~ \input{images/CellSmall}
            \caption[Sketch how the Cell Method works]
            {
                Sketch how the cell algorithm for the construction of the
                RNG and GG works. Here with the lune definition of the GG.
                The bounding box of the lune, which determines which cells have
                to be tested, is marked with thick colored lines.
                Left: shows that it is sufficient to find a
                single node in an inner cell of the bounding box to discard
                the edge between two distant nodes. The nodes inside the
                other 8 marked cells do not have to be tested anymore.
                Right: shows that only the nodes inside the
                marked cell have to be tested, because there are no other nodes,
                the edge can be drawn.
            }
            \label{fig:cell}
        \end{figure}\\
        Given the discretized cell structure it is just necessary to test the nodes in the cells which
        resemble a rectangular bounding box of the lune. Most pairs will be
        far away from each other and there will be one or more cells in the
        middle of the bounding box, which are completely inside the lune.
        Then only one node has to be tested from such a cell to discard an
        edge between the nodes. On the other hand, nodes that will be connected
        are near to each other so that only very few cells intersect the bounding
        box of their lune and consequently only very few nodes have to be tested.\\
        %~ Here the number of cells equals the number of nodes.
        The complexity of this method should be between \(O(N^2)\) in the
        best case, because for every pair at least one check has to be performed,
        and \(O(N^3)\) in the worst case, e.g.\ all nodes are inside one
        cell.\\
        Combining both algorithms yields a blazing fast algorithm in complexity
        \(O(N \log N)\) for nodes which are evenly spread across the cells.\\
        However, generation of the graphs is not time critical, because
        the MC simulations take much longer. So only the cell algorithm
        was used to gegnerate the graphs.\\
